---
title: "Profiles and positions analysis"
author: "Rydberg"
date: "25 de febrero de 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
packages <- c('dplyr', 'caret', 'olsrr', 'car', 'rgl','reshape2','ggplot2','biobroom','tidyr',
              'dendextend','matrixStats','cluster','gridExtra','grid','lattice')
for(i in packages){library(package=i,character.only = T)}
```

## Figure 1

Loading data sets of fluctuations, electrostatic potentials and rmsf. Additionally load kinetic paramters

```{r}
# Fluctuations dataset
data_nma <- read.table('fluctuations.csv ', sep=',', header = T, row.names = 1 )/10
colnames(data_nma) <-1:185
# Electrostatic potentials dataset
data_ep <- read.table('electrostatic_potentials.csv ', sep=';', header = T, row.names = 1 )
colnames(data_ep) <-1:185
# RMSF backbone dataset
data_md_b <- read.table('rmsf_backbone.csv ', sep=';', header = T, row.names = 1 )
colnames(data_md_b) <-1:185
# kinetic parameters dataset
data_k <- read.table('kinetic_parameters.csv ', sep=';', header = T, row.names = 1 )
```

Transpose each data set and add a variable identifying each position

```{r}
# Define position variable
Position <- 1:185
# Transpose fluctuations dataset
subdata_nma <- as.data.frame(t(data_nma))
subdata_nma <- cbind(subdata_nma,Position)
# Transpose electrostatic potentials dataset
subdata_ep <- as.data.frame(t(data_ep))
subdata_ep <- cbind(subdata_ep,Position)
# Transpose RMSF backbone dataset
subdata_md_b <- as.data.frame(t(data_md_b))
subdata_md_b <- cbind(subdata_md_b,Position)
```

Convert wide to long datasets by gathering the measurements by positions and using the name of the mutation as a key. Exclude position from the gathering to conserve it. 

The final output should be a dataset with three columns, describing the mutation, position, and the measurement of interest.

```{r}
# Gather fluctuations dataset and convert it to long format
subdata_nma <- gather(subdata_nma,key='Mutation',value='Fluctuation',-Position)
# Gather electrostatic potentials dataset and convert it to long format
subdata_ep <- gather(subdata_ep,key='Mutation',value='DEP',-Position)
# Gather RMSF backbone dataset and convert it to long format
subdata_md_b <- gather(subdata_md_b,key='Mutation',value='RMSF',-Position)
```

Create the ggplots objects

```{r}
# Define the theme
theme_plots = theme(axis.text = element_text(size=12), axis.title = element_text(size=10),
                    legend.text = element_text(size=12),
                    legend.title = element_text(size=12),
                    legend.position = 'none',
                    axis.line = element_line(colour = "black"), 
                    panel.background = element_blank(), panel.grid.major = element_blank(),
                    panel.grid.minor = element_blank())
# Plot 1: fluctuations
g1 <- ggplot(subdata_nma,mapping = aes(x=Position,y=Fluctuation,group=Mutation)) + 
  geom_line(size=0.7, color=alpha('grey',0.7)) + 
  geom_line(size=0.7, aes(color=Mutation), data=subdata_nma[subdata_nma['Mutation']=='WT',],
            color = 'red') + theme_plots +
  scale_x_continuous(breaks=seq(1,185,23), limits = c(1,185)) + 
  ylab('Fluctuation (nm)')
# Plot 2: electrostatic potentials
g2 <- ggplot(subdata_ep,mapping = aes(x=Position,y=DEP,group=Mutation)) + 
  geom_line(size=0.7, color=alpha('grey',0.7)) + 
  geom_line(size=0.7, aes(color=Mutation), data=subdata_ep[subdata_ep['Mutation']=='WT',],
            color = 'red') +
  theme_plots +
  scale_x_continuous(breaks=seq(1,185,23), limits = c(1,185)) + 
  ylab('Difference in Electrostatic Potential (mV)')
# Plot 3: RMSF backbone
g3 <- ggplot(subdata_md_b,mapping = aes(x=Position,y=RMSF,group=Mutation)) + 
  geom_line(size=0.7, color=alpha('grey',0.7)) + 
  geom_line(size=0.7, aes(color=Mutation),
            data=subdata_md_b[subdata_md_b['Mutation']=='WT',], color = 'red') +
  theme_plots +
  scale_x_continuous(breaks=seq(1,185,23), limits = c(1,185)) + 
  ylab('RMSF backbone (nm)')
```

Save the figures as figure 1

```{r}
grob <- arrangeGrob(g1, g3, g2)
ggsave('fig1.tiff',grob,width = 8, height = 10, dpi = 300)
```

## Supplementary figures

Bind each dataset with the kinetic data

```{r}
# Bind fluctuation dataset to kcat data
data_nma_k <- cbind(data_k[1],data_nma)
# Bind electrostatic potential dataset to kcat data
data_ep_k <- cbind(data_k[1],data_ep)

# Bind RMSF backbone dataset to kcat data
data_md_b_k <- cbind(data_k[1],data_md_b)

# Bind RMSF protein dataset to kcat data
# data_md_p_k <- cbind(data_k[1],data_md_p)
```

Create a new variable that indicates if relative kcat is greater than 50

```{r}
# Define a kcat cutoff variable
kcat_cutoff = 50
cutoff=as.factor(ifelse(data_k[,1]>=kcat_cutoff, paste('\u2265',kcat_cutoff, sep=''),
                             paste('< ',kcat_cutoff, sep='')))

# Add the kcat cutoff variable to kinetic fluctuation dataset
data_nma_g <- mutate(.data=data_nma_k, Kcat = cutoff)

# Add the kcat cutoff variable to kinetic electrostatic potential dataset
data_ep_g <- mutate(.data=data_ep_k, Kcat = cutoff)

# Add the kcat cutoff variable to kinetic RMSF backbone dataset
data_md_b_g <- mutate(.data=data_md_b_k, Kcat = cutoff)

# Add the kcat cutoff variable to kinetic RMSF protein dataset
# data_md_p_g <- mutate(.data=data_md_p_k, Kcat = cutoff)
```

Boxplot for the whole fluctuation profile for each group of PZAses 

```{r}
# Define a kcat cutoff variable
kcat_cutoff = 50
cutoff=as.factor(ifelse(data_k[,1]>=kcat_cutoff, paste('\u2265',kcat_cutoff, sep=''),
                             paste('< ',kcat_cutoff, sep='')))
# Create a list with the dataframes to evaluate and another list with the y-axis titles for each subplot.
L = list(data_nma_k, data_ep_k, data_md_b_k)
D = list('Fluctuations (nm)', 'Difference in Electrostatic Potentials (mV)', 'RMSF backbone (nm)')
# Create a figure lists
figs_s = list()
figs_r = list()
# Generate individual figures
for(i in 1:length(L)){
  
  label = D[[i]]
  dataset = L[[i]]
  
  data_L_r <- filter(dataset, Kcat < kcat_cutoff)
  data_L_r <- mutate(.data=data_L_r,
                   Kcat= as.factor(ifelse(data_L_r[,1]>=kcat_cutoff,
                                          paste('\u2265',kcat_cutoff, sep=''), 
                                          paste('< ',kcat_cutoff, sep=''))))
  
  data_L_s <- filter(dataset, Kcat >= kcat_cutoff)
  data_L_s <- mutate(.data=data_L_s, 
                     Kcat = as.factor(ifelse(data_L_s[,1]>=kcat_cutoff,
                                             paste('\u2265',kcat_cutoff, sep=''),
                                             paste('< ',kcat_cutoff, sep=''))))
  
  bp_r <- gather(data_L_r,key='Position',value='Descriptor',-Kcat,factor_key = T)
  bp_s <- gather(data_L_s ,key='Position',value='Descriptor',-Kcat,factor_key = T)
  
  
  theme_plots = theme(axis.text = element_text(size=18), axis.title = element_text(size=18),
                    legend.text = element_text(size=12),
                    legend.title = element_text(size=12),
                    axis.line = element_line(colour = "black"), 
                    panel.background = element_blank(), panel.grid.major = element_blank(),
                    panel.grid.minor = element_blank())
  g1 <- ggplot(bp_s,aes(x=Position,y=Descriptor, fill=Kcat)) + 
    stat_boxplot(geom='errorbar') +
    geom_boxplot(outlier.shape = 21, fill='#00BFC4', outlier.size=0.8, size = 0.1) +
    theme_plots + scale_x_discrete(breaks=seq(1,185,23), limits = 1:190) + ylab(label) +
    ylim(min(dataset[-1]),max(dataset[-1]))
  g2 <- ggplot(bp_r,aes(x=Position,y=Descriptor, fill=Kcat)) +
    stat_boxplot(geom='errorbar') +
    geom_boxplot(outlier.shape = 21, fill='#F8766D', outlier.size=0.8, size = 0.1) +
    theme_plots + scale_x_discrete(breaks=seq(1,185,23), limits = 1:190) +  ylab(label) +
    ylim(min(dataset[-1]),max(dataset[-1]))
  
  figs_s[[i]] <- g1 
  figs_r[[i]] <- g2
}
# Create and save the figure
grob <- arrangeGrob(figs_s[[1]], figs_r[[1]], figs_s[[3]], figs_r[[3]], figs_s[[2]], figs_r[[2]])
ggsave('figsup1.tiff' ,grob,width = 16, height = 17, dpi = 300)
  
```

## Statistical tests

Perform a non-parametric statistical test for position between the two groups of PZAses for fluctuations, RMSFs, and electrostatic potentials datasets.

```{r, warning=FALSE,message=FALSE, error=FALSE }
# Identity of each residue in PZAse
PZAse <-c('MRALIIVDVQNDFCEGGSLAVTGGAALARAISDYLAEAADYHHVVATKDFHIDPGDHFSGTPDYSSSWPPHCVSGTPGADFHPSLDTSAIEAVFYKGAYTGAYSGFEGVDENGTPLLNWLRQRGVDEVDVVGIATDHCVRQTAEDAVRNGLATRVLVDLTAGVSADTTVAALEEMRTASVELVCS')
PZAse <- paste(strsplit(PZAse,'')[[1]],1:185,sep='')
# Name dataframes
colnames(data_nma_g) <- c('Kcat',paste('x',1:185,sep=''))
colnames(data_ep_g) <- c('Kcat',paste('x',1:185,sep=''))
colnames(data_md_b_g) <- c('Kcat',paste('x',1:185,sep=''))
# Define a list for each dataset to evaluate and a list for the name of the csv with the results.
L = list(data_nma_g, data_ep_g, data_md_b_g)
N = list('Test_nma.csv','Test_ep.csv','Test_md_b.csv')
# Iterate and perform the non-parametric test for each dataset
for(n in 1:length(L)){
  
  # Select the dataset and create storing variables
  dataset <- L[[n]]
  dataset[-1] <- dataset[-1]
  position <- c()
  p.values <- c()
  conf.int.lower <- c()
  conf.int.upper <- c()
  # Actually perform the statistical test, record intervals and p-values
  for(i in colnames(dataset[-1])){
    test <- wilcox.test(as.formula(paste(i,"~Kcat",sep="")),
                        alternative = 'two.sided',dataset,conf.int=T)
    position <- c(position,i)
    conf.int.lower <- c(conf.int.lower,test$conf.int[1])
    conf.int.upper <- c(conf.int.upper,test$conf.int[2])
    p.values <- c(p.values,test$p.value)
  }
  
  # Calculate adjusted p-balues
  adj.p.values <- p.adjust(p.values, method ='BH')
  
  # Create and save the csv files with the results
  test.table <-  cbind(position,conf.int.lower,conf.int.upper,p.values,adj.p.values)
  write.table(test.table,N[[n]],sep=',',quote = F, row.names = F)
  # Print positions that are called significant with and without multiple testing correction
  print(c(N[[n]],PZAse[as.numeric(test.table[,'p.values'])<0.05]))
  print(c(N[[n]],PZAse[as.numeric(test.table[,'adj.p.values'<0.05])],' adjusted'))
}
```
